Finite Property Convex Elicitation Paper

### 2019-01-25 cvx-fin.tex
maybe half done with 1d proofs

\newcommand{\prop}[1]{\mathsf{prop}[#1]}
\newcommand{\eliccvx}{\mathrm{elic}_\mathrm{cvx}}
\newcommand{\elicpoly}{\mathrm{elic}_\mathrm{pcvx}}
\newcommand{\elicembed}{\mathrm{elic}_\mathrm{embed}}

@@

\subsection{Notation}
\begin{itemize}
	\item $\Y$ is the finite outcome space, and $n := |\Y|$.
	\item $\simplex$ is the set of distributions over $\Y$.
	\item For an outcome $y$, we denote $p_y$ as the probability of outcome $y$ being observed in nature.
	\item If $\gamma = \prop{\ell}$, then $\gamma:\simplex \toto \R$ is the property elicited by the loss $\ell: \R \to \reals^\Y$.
	\item $d$ is the dimension of the embedding space
	\item $L:\reals^d \to \reals^\Y$ is a loss taking a report $u \in \reals^d$ and mapping the loss over each outcome in $\Y$.
	\item $\Gamma: \simplex \toto \reals^d$ is the set-valued property mapping a probability distribution to a report $u \in \reals^d$.
	\item $\Gamma_u = \{ p \in \simplex : u \in \Gamma(p) \}$
	\item $\psi$ is the link function (see below)
	\item $\varphi:\R \to \reals^d$ is the embedding function
	\item $\trim(\Gamma)$ and $\strip(\Gamma)$
	\item $\eliccvx$, $\elicembed$, and $\elicpoly$
	\item $V$ and $B(P)$ as we get ready for later.
\end{itemize}

@@

@@

\begin{definition}
  Loss $L:\R\to\reals^\Y$, elicits
\end{definition}

@@

\begin{definition}\label{def:strip-and-trim}
	\jessie{strip, trim}
	For $\Gamma:\simplex \toto \R'$, define $\strip(\Gamma) = \{\Gamma_r : r \in \R' \}$.
	If $\Gamma$ has a finite set of full-dimensional level sets that union to $\simplex$, there is a finite property $\gamma:\simplex \toto \R$ such that $\strip(\gamma) := \trim(\Gamma)$.
	\jessie{Fix without over-explaining.}
\end{definition}

\begin{proposition}\label{prop:optimal-reports-per-level-set}
	Let $\Gamma:\simplex\toto\reals^d$ be a non-degenerate (convex) elicitable property.

	The following are equivalent:
	\begin{enumerate}
		\item There is a nondegenerate, finite property $\gamma:\simplex\toto\R$ such that there is an injection $\varphi:\R\to\reals^d$ such that $r\in\gamma(p) \iff \varphi(r) \in \Gamma(p)$. (i.e. $\Gamma$ embeds a finite $\gamma$).  
		\item $\trim(\Gamma)$ is finite.     
		\item There is a finite set of full-dimensional level sets $\Theta$ of $\Gamma$ that union to $\simplex$.

	\end{enumerate}
\end{proposition}

\begin{proof}
	We proceed in a cycle to show $1 \implies 2 \implies 3 \implies 1 $.
	\begin{enumerate}
		\item [$1 \implies 2$]
		We know $\strip(\gamma) = \trim(\Gamma)$ (since $\varphi$ is an injection, and there is some intermediate property $\Gamma'$ such that $\Gamma' \preceq \Gamma$ and with a bijection $\phi:\Gamma'(p) \mapsto \gamma(p)$, so $\strip(\Gamma') = \strip(\gamma)$.) 
		As $\gamma$ is finite, so is $\strip(\gamma)$ since strip is generated from the (finite) report set.

		\item [$2 \implies 3$]
		Since $\Gamma$ is nondegenerate and every $\theta \in \trim(\Gamma)$ is full dimensional, it is clear that the union of the level sets of $\trim(\Gamma)$ union to $\simplex$ by nondegneracy of $\Gamma$.

		\item[$3 \implies 1$]
		Construct $\R := \{[u] \in \Gamma(p) : p \in \inter{\theta} \}_{\theta \in \Theta}$.
		The property $\gamma: p \mapsto \Gamma(p) \cap\R$ is then finite as $\R$ is finite.
		Taking $\varphi$ to be the identity, $\Gamma$ embeds this finite $\gamma$.
	\end{enumerate} 
\end{proof}

@@

\begin{definition}
  A property $\Gamma:\simplex\to\R$ is \emph{monotone} if there are maps $a:\R\to\reals^\Y$, $b:\R\to\reals^\Y$ and a total ordering $<$ of $\R$ such that the following two conditions hold.
  \begin{enumerate}
  \item For all $r\in\R$, we have $\Gamma_r = \{p\in\simplex : a(r) \cdot p \leq 0 \leq b(r) \cdot p\}$.
  \item For all $r < r'$, we have $b(r) \leq a(r')$ (component-wise).
  % \item For all $r,r'\in\R$ and $p\in\Gamma_{r'}\setminus\Gamma_r$, we have $b(r) \cdot p < 0 \implies r' > r$ and $a(r) \cdot p > 0 \implies r' < r$.
  \end{enumerate}
\end{definition}

\raf{From Lambert et al. (2009) -- or maybe this is only in his unpublished paper; need to check}
\begin{definition}
  A finite property $\gamma:\simplex\to\R$ is \emph{orderable} if there is an enumeration $\R = \{r_1,\ldots,r_k\}$ such that for all $i\in\{1,\ldots,k-1\}$, the intersection $\gamma_{r_i} \cap \gamma_{r_{i+1}}$ is a hyperplane intersected with $\simplex$.
\end{definition}

\begin{lemma}
  A finite property is orderable if and only if it is monotone.
\end{lemma}

\begin{lemma}
  For any convex loss $L : \reals \to \reals^\Y$, $\prop{L}$ is monotone.
\end{lemma}

\begin{proposition}
  If convex $L : \reals \to \reals^\Y$ indirectly elicits a finite elicitable property $\gamma$, then $\gamma$ is orderable.
\end{proposition}

\raf{Now follows from the previous and polyhedral $\implies$ embeds}
\begin{corollary}
  Every polyhedral $L : \reals \to \reals^\Y$ embeds an orderable property.
\end{corollary}

\begin{proposition}
  Every orderable property $\gamma$ is 1-embeddable.
\end{proposition}

>>>>

\newcommand{\prop}[1]{\Gamma[#1]}
\newcommand{\eliccvx}{\mathrm{elic}_\mathrm{cvx}}
\newcommand{\elicpoly}{\mathrm{elic}_\mathrm{pcvx}}
\newcommand{\elicembed}{\mathrm{elic}_\mathrm{embed}}

@@

\subsection{Notation}
\begin{itemize}
	\item $\Y$ is the finite outcome space, and $n := |\Y|$.
	\item $\simplex\subseteq\reals^\Y$ is the set of distributions over $\Y$, thought of as vectors of probabilities.
	\item For an outcome $y$, we denote $p_y$ as the probability of outcome $y$ being observed in nature.
	\item If $\gamma = \prop{\ell}$, then $\gamma:\simplex \toto \R$ is the property elicited by the loss $\ell: \R \to \reals^\Y$.
	\item $d$ is the dimension of the embedding space
	\item $L:\reals^d \to \reals^\Y$ is a loss taking a report $u \in \reals^d$ and mapping the loss over each outcome in $\Y$. \raf{Changed to $\reals^\Y_+$ below.}
	\item $\Gamma: \simplex \toto \reals^d$ is the set-valued property mapping a probability distribution to a report $u \in \reals^d$.
	\item $\Gamma_u = \{ p \in \simplex : u \in \Gamma(p) \}$
	\item $\psi$ is the link function (see below)
	\item $\varphi:\R \to \reals^d$ is the embedding function
	\item $\trim(\Gamma)$ and $\strip(\Gamma)$
	\item $\eliccvx$, $\elicembed$, and $\elicpoly$
	\item $V$ and $B(P)$ as we get ready for later.
\end{itemize}

@@

  Level set $\Gamma_r$

\raf{Should bake non-degenerate into the definition: insist that $\Gamma(p) \neq \emptyset$ for all $p$.}

@@

\begin{definition}
  \raf{I realized that (a) it is common to restrict losses to be non-negative, and (b) it solves some annoying issues where we have to say ``if $L$ is bounded below...'' nicely.  So unless any objections, let's say $L \geq 0$.}

  Loss $L:\R\to\reals^\Y_+$, elicits
\end{definition}

@@

% \begin{definition}\label{def:strip-and-trim}
% 	\jessie{strip, trim}
% 	For $\Gamma:\simplex \toto \R'$, define $\strip(\Gamma) = \{\Gamma_r : r \in \R' \}$.
% 	If $\Gamma$ has a finite set of full-dimensional level sets that union to $\simplex$, there is a finite property $\gamma:\simplex \toto \R$ such that $\strip(\gamma) := \trim(\Gamma)$.
% 	\jessie{Fix without over-explaining.}
% \end{definition}

\begin{proposition}\label{prop:optimal-reports-per-level-set}
  \raft{Jessie: I think this should be fine for the paper.  We may want strip and trim for the proof, in the appendix, in which case we can define them there...}
	Let $\Gamma:\simplex\toto\reals^d$ be a non-degenerate (convex) elicitable property.
    \raf{Do we really need $\Gamma$ to be convex elicitable?}

	The following are equivalent:
	\begin{enumerate}
		\item $\Gamma$ embeds a finite elicitable property.
		\item The set $\{\Gamma_r : r \in \R' \}$ is finite.
		\item There is a finite set of full-dimensional level sets $\Theta$ of $\Gamma$ that union to $\simplex$.
	\end{enumerate}
\end{proposition}

% \begin{proof}
% 	We proceed in a cycle to show $1 \implies 2 \implies 3 \implies 1 $.
% 	\begin{enumerate}
% 		\item [$1 \implies 2$]
% 		We know $\strip(\gamma) = \trim(\Gamma)$ (since $\varphi$ is an injection, and there is some intermediate property $\Gamma'$ such that $\Gamma' \preceq \Gamma$ and with a bijection $\phi:\Gamma'(p) \mapsto \gamma(p)$, so $\strip(\Gamma') = \strip(\gamma)$.) 
% 		As $\gamma$ is finite, so is $\strip(\gamma)$ since strip is generated from the (finite) report set.

% 		\item [$2 \implies 3$]
% 		Since $\Gamma$ is nondegenerate and every $\theta \in \trim(\Gamma)$ is full dimensional, it is clear that the union of the level sets of $\trim(\Gamma)$ union to $\simplex$ by nondegneracy of $\Gamma$.

% 		\item[$3 \implies 1$]
% 		Construct $\R := \{[u] \in \Gamma(p) : p \in \inter{\theta} \}_{\theta \in \Theta}$.
% 		The property $\gamma: p \mapsto \Gamma(p) \cap\R$ is then finite as $\R$ is finite.
% 		Taking $\varphi$ to be the identity, $\Gamma$ embeds this finite $\gamma$.
% 	\end{enumerate} 
% \end{proof}

@@

\begin{definition}
  A property $\Gamma:\simplex\to\R$ is \emph{monotone} if there are maps $a:\R\to\reals^\Y$, $b:\R\to\reals^\Y$ and a total ordering $<$ of $\R$ such that the following two conditions hold.
  \begin{enumerate}
  \item For all $r\in\R$, we have $\Gamma_r = \{p\in\simplex : a(r) \cdot p \leq 0 \leq b(r) \cdot p\}$.
  \item For all $r < r'$, we have $a(r) \leq b(r) \leq a(r') \leq b(r')$ (component-wise).
  % \item For all $r,r'\in\R$ and $p\in\Gamma_{r'}\setminus\Gamma_r$, we have $b(r) \cdot p < 0 \implies r' > r$ and $a(r) \cdot p > 0 \implies r' < r$.
  \end{enumerate}
\end{definition}

\raf{From Lambert et al. (2009) -- or maybe this is only in his unpublished paper; need to check}
\begin{definition}
  A finite property $\gamma:\simplex\toto\R$ is \emph{orderable} if there is an enumeration $\R = \{r_1,\ldots,r_k\}$ such that for all $i\in\{1,\ldots,k-1\}$, the intersection $\gamma_{r_i} \cap \gamma_{r_{i+1}}$ is a hyperplane intersected with $\simplex$.
\end{definition}

\begin{lemma}\label{lem:orderable-monotone}
  A finite property is orderable if and only if it is monotone.
\end{lemma}
\begin{proof}
  Let $\gamma:\simplex\toto\R$ be finite and orderable.
  For $1 \leq i < k$, let $v_i\in\reals^\Y$ such that $\gamma_{r_i} \cap \gamma_{r_{i+1}} = \{p\in\simplex : v_i\cdot p = 0\}$.
  As we assume $\gamma$ is non-redundant, we have $p^1 \in \inter\gamma_{r_1}$.
  Without loss of generality, we take the $v_i$ above so that $v_i \cdot p^1 > 0$ for all $i$.
  Now let $b(1) = v_1$ and $a(1) = -\max_y |b(1)_y|\ones$.
  \raf{Really not sure if that is the right start.}
  By our assumptions, $p \in \gamma_1 \iff b(1) \cdot p \geq 0$.
  \raf{Show how to make them monotone... I remember doing this for the continuous convex paper...}

  Conversely, suppose $\gamma:\simplex\toto\R$ be finite and monotone.
  Then we can use the total ordering of $\R$ to write $\R = \{r_1,\ldots,r_k\}$ such that $r_i < r_{i+1}$ for all $i \in \{1,\ldots,k-1\}$.
  We now have $\gamma_{r_i} \cap \gamma_{r_{i+1}} = \{p\in\simplex : a(r_{i+1}) \cdot p \leq 0 \leq b(r_i) \cdot p\}$.
  If this intersection is empty, then there must be some $p$ with $b(r_i) \cdot p < 0$ and $a(r_{i+1}) \cdot p > 0$; by monotonicity, no earlier or later reports can be in $\gamma(p)$, so we see that $\gamma(p) = \emptyset$, a contradiction.
  Thus the intersection is nonempty, and as we also know $b(r_i) \leq a(r_{i+1})$ we conclude $b(r_i) = a(r_{i+1})$, and the intersection is the hyperplane defined by $b(r_i) = a(r_{i+1})$.
\end{proof}

\begin{lemma}\label{lem:prop-L-monotone}
  For any convex loss $L : \reals \to \reals^\Y$, $\prop{L}$ is monotone.
\end{lemma}
\begin{proof}
  Let $a,b$ be defined by $a(r)_y = \partial_- L(r)_y$ and $b(r) = \partial_+ L(r)_y$, that is, the left and right derivatives of $L(\cdot)_y$ at $r$, respectively.
  Then $\partial L(r)_y = [a(r)_y,b(r)_y]$.
  We now have $r \in \prop{L}(p) \iff 0 \in \partial p\cdot L(r) \iff a(r)\cdot p \leq 0 \leq b(r) \cdot p$, showing the first condition.
  The second condition follows as the subgradients of $L$ are monotone functions.
  \raf{More detail and some Rockafellar cites}
\end{proof}

\begin{proposition}\label{prop:indirect-orderable}
  If convex $L : \reals \to \reals^\Y$ indirectly elicits a finite elicitable property $\gamma$, then $\gamma$ is orderable.
\end{proposition}
\begin{proof}
  Let $\gamma:\simplex\toto\R$.
  From Lemma~\ref{lem:prop-L-monotone}, $\prop{L}$ is monotone.
  Let $\psi:\reals\to\R$ be the calibrated link from $\prop{L}$ to $\gamma$.

  \raf{FROM IPAD NOTES}
\end{proof}

\raf{Now follows from the previous and polyhedral $\implies$ embeds}
\begin{corollary}\label{cor:embed-orderable}
  Every polyhedral $L : \reals \to \reals^\Y$ embeds an orderable property.
\end{corollary}

\begin{proposition}
  Every orderable property $\gamma$ is 1-embeddable.
\end{proposition}
\begin{proof}
  Let $\R = \{r_1,\ldots,r_k\}$ be the report space for $\gamma$.
  From Lemma~\ref{lem:prop-L-monotone}, and its proof, we have $a:\R\to\reals^\Y, b\to\reals^\Y$ such that $\gamma_r = \{p\in\simplex : a(r) \cdot p \leq 0 \leq b(r) \cdot p\}$ for all $r\in\R$, and for all $i \in \{1,\ldots,k-1\}$, we have $a(r_i) \leq b(r_i) = a(r_{i+1}) \leq b(r_{i+1})$.
  We now define $\varphi(r_i) = i \in \reals$, and $L$... \raf{integrate?  this should be simple... want $\partial L(i)_y = [a(r_i)_y,b(r_i)_y]$.}
\end{proof}
